Decoding Toronto's Crime Patterns

### 2024-01-25 outputs/paper/paper.qmd
Improve writing

---
title: "Decoding Toronto's Crime Patterns"
subtitle: "A Comprehensive Analysis of Victims and Trends (2014-2022)"
author: "Chay Park"
date: today
date-format: long
abstract: "In the period from 2014 to 2022, including the global Novel Coronavirus pandemic in 2019, this paper scrutinizes the Toronto Police annual statistical report on victims of crimes. Analyzing this extensive dataset reveals a correlation between the pandemic and crime rates, emphasizing broader trends in the top three crimes across different age groups. The findings underscore the nuanced relationship between external events, law enforcement, and crime reporting, offering insights into the complex interplay between specific events and crime statistics. This research contributes to a comprehensive understanding of societal dynamics, highlighting the potential impact of the pandemic on both crime rates and the reporting of victims of crimes over the years."
format: pdf
prefer-html: true
toc: true
toc-depth: 3
toc-title: Contents
thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Toronto_CrimeVictims."
number-sections: true
number-depth: 3
bibliography: references.bib
---

@@

Analyzing this dataset becomes imperative as it unravels a nuanced narrative beneath the overarching claim of declining crime rates in Canada, as reported by Statistics Canada. Contrary to the general trend, the Statista Research department asserts that the overall crime rate in Canada has consistently decreased since 2000, with the 2021 crime rate approximately 30 percent lower than its peak in 2003. Property crime aligns with this decline, dropping by 30.6 percent during the same period (Statista Research Department 2021). However, a surprising surge in violent crime rates in 2021 challenges this narrative, suggesting a potential consequence of crime rates rebounding after a decline during the pandemic. This analysis focuses on discerning general trends in victims reporting specific crimes---robbery, assault, and sexual violation---utilizing tables and graphs to visualize these trends from 2014 to 2022. The objective is to identify distinctive patterns within the trend, considering variations across age groups and different crime types. Additionally, an exploration of potential correlations with external factors, such as the global pandemic, will be conducted.

This paper employs R (R Core Team 2020) to analyze a dataset sourced from opendatatoronto (Gelfand 2020), encompassing all identified victims of crimes against persons. The dataset under analysis comprises observations on the number of victims reported each year for crimes, including robbery, assault, and sexual violation, to the Toronto Police from 2014 to 2022. This inclusive dataset considers cases that may have been deemed unfounded post-investigation, incidents occurring beyond the City of Toronto limits, or those lacking verified locations. The analysis extracts key variables such as report_year, subtype, age_group, and count\_, excluding age_cohort, sex, category, and assault_subtype. Omitted are unknown values for age_group, age_cohort, and sex, as they don't contribute explicitly to defining data groups. The dataset lacks explicit details regarding these unknown values. Additionally, the analysis focuses solely on the top three reported crimes---robbery, assault, and sexual violation. The exclusion of 'other' crimes in the cleaned_data set is attributed to its inherent uncertainty. Notably, the data within this paper centers on trends related to these three crimes, with age groups classified into three subgroups: adult, youth, and child. The overarching trend indicates a resurgence in reported victims of crimes following a period of decline, 2020 - 2021.

The paper begins with a comprehensive overview, presenting a table illustrating the total number of victims across various age groups over the specified period. This overview is complemented by corresponding graphs derived from the table. The subsequent section delves into bar graphs detailing the reported victims' numbers within distinct age groups, individually addressing sexual violation, assault, and robbery. Finally, a concise table is presented, offering a quantitative measure of the number of victims for each crime across the report years.

@@

@tbl-totalsum-table above illustrates the overarching trend in the number of victims across distinct age groups. The age categorization is determined by the reporting year, with the "Adult" group encompassing individuals aged 18 or older. The "Youth" group comprises individuals between the ages of 12 and 17, while those below 12 fall into the "Child" group. This presentation aims to provide a comprehensive overview of victim counts within these age cohorts over the years from 2014 to 2022.

@@

@tbl-totalsum-table counts the total number of victims for each age group over the years. It is evident that the "Adult" victim group has consistently reported the highest number of incidents, while the "Child" victim group reports the least. Notably, the year 2020 stands out, revealing a substantial reduction in reported victim counts across all three age groups. This decline is attributed to the impact of COVID-19 and the ensuing lockdown measures, resulting in a significant reduction in crime reporting rates.

@@

@fig-totalsum-graph corresponds to the information presented in @tbl-totalsum-table. Visualizing the data by age group across each year reveals that the majority of reported crime victims fall into the "Adult" category. Notably, the bars exhibit an ascending trend from 2014 to 2018. However, a conspicuous drop in reported crimes is observed between 2019 and 2020. One plausible explanation for this decline is attributed to the impact of COVID-19, which significantly influenced crime reporting rates during this period.

@@

@fig-r-graph depicts that as of 2018, the youth group of robbery victims diminished and has remained at a lower level. The COVID pandemic is a significant factor influencing the bar graphs depicting the number of robbery victims in different age groups over the years. In 2020, victims of robbery in every age group significantly decreased, indicating a possible impact of lockdown mandates on the decrease in victims.

@@

The onset of the COVID-19 pandemic in late 2019 introduced a significant factor that impacted crime trends throughout the subsequent years, especially in 2020 and extending up to 2022. One plausible explanation is the reduction in physical interactions among individuals, potentially leading to a decline in contact-based offenses such as sexual violence and robbery. However, it is crucial not to overlook the possibility of an increase in domestic violence during this period. Domestic settings may contribute to instances of sexual violence, presenting a contrasting outcome.

@@

The analysis reveals notable trends in victims of crimes in Toronto from 2014 to 2022. While the overall pattern suggests a decline in reported crimes, the impact of the COVID-19 pandemic is evident in specific crime subtypes and age groups. Sexual violation cases, for instance, witnessed a shift in the dominant age group from adults to youth in 2021. Assault incidents displayed a consistent upward trend, unaffected by the pandemic, but with a reduction in the sizes of child and youth victim groups. In contrast, the number of robbery victims, particularly in the youth group, significantly decreased in 2020, likely influenced by pandemic-related lockdown measures. These findings provide valuable insights into the complex dynamics of crime reporting and its interaction with external events.

>>>>

---
title: "Decoding Toronto's Crime Patterns"
subtitle: "A Comprehensive Analysis of Victims and Trends (2014-2022)"
author: "Chay Park"
date: today
date-format: long
abstract: "In the period from 2014 to 2022, including the global Novel Coronavirus pandemic in 2019, this paper scrutinizes the Toronto Police annual statistical report on victims of crimes. By analyzing trends in the top three crimes across different age groups, the paper conveys a message on a correlation between the pandemic and crime rates. This research contributes to an understanding of societal dynamics, highlighting the potential impact of the pandemic on both crime rates and the reporting of victims of crimes over the years."
format: pdf
prefer-html: true
toc: true
toc-depth: 3
toc-title: Contents
thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Toronto_CrimeVictims."
number-sections: true
number-depth: 3
bibliography: references.bib
---

@@

According to U.S. News, Canada placed second in the overall ranking for the best countries in 2023. One of the highest-scoring categories, 'Quality of Life,' ranks third, and its attribute, 'Safe,' scored 93.2 out of 100 (Best Countries 2023). Indeed, the Statista Research Department asserts that the overall crime rate in Canada has consistently decreased since 2000, with the 2021 crime rate approximately 30 percent lower than its peak in 2003 (Statista Research Department 2021). However, a surge in violent crime rates in 2021 challenges this claim, suggesting a potential consequence of crime rates rebounding after a decline during the pandemic. This analysis focuses on discerning general trends in victims reporting specific crimes---robbery, assault, and sexual violation---utilizing tables and graphs to visualize these trends from 2014 to 2022. The objective is to identify distinctive patterns within the trend, considering variations across age groups and different crime types. Additionally, an exploration of potential correlations with external factors, such as the global pandemic, will be conducted.

This paper employs R (R Core Team 2020) to analyze a dataset sourced from opendatatoronto (Gelfand 2020), encompassing all identified victims of crimes against persons. The dataset comprises observations on the number of victims reported each year to the Toronto Police from 2014 to 2022. This considers cases that may have been deemed unfounded post-investigation, incidents occurring beyond the City of Toronto limits, or those lacking verified locations (Toronto Police Services 2023). The analysis extracts variables such as report_year, subtype, age_group, and count\_, excluding age_cohort, sex, category, and assault_subtype. Omitted are unknown values for age_group, age_cohort, and sex, as they do not contribute explicitly to defining data groups. The dataset lacks details regarding these unknown values. Additionally, the analysis focuses on the top three reported crimes---robbery, assault, and sexual violation. The exclusion of 'other' crimes in the cleaned_data set is attributed to its uncertainty. The data within this paper center on trends related to these three crimes, with age groups classified into three subgroups: adult, youth, and child. The trend indicates a resurgence in reported victims of crimes following a period of decline, 2020-2021.

The paper begins with a general overview of the dataset, presenting a table illustrating the total number of victims across various age groups over the specified period. This overview is complemented by corresponding graphs. The subsequent section delves into bar graphs detailing the reported victims' numbers within distinct age groups, individually addressing sexual violation, assault, and robbery. Finally, a table offers a quantitative measure of the number of victims for each crime across the report years.

@@

@tbl-totalsum-table above illustrates the trend in the number of victims across distinct age groups. The age categorization is determined by the reporting year, with the "Adult" group encompassing individuals aged 18 or older. The "Youth" group comprises individuals between the ages of 12 and 17, while those below 12 fall into the "Child" group. This presentation aims to provide a comprehensive overview of victim counts within these age cohorts over the years from 2014 to 2022.

@@

@tbl-totalsum-table counts the total number of victims for each age group over the years. It is evident that the "Adult" victim group has consistently reported the highest number of incidents, while the "Child" victim group reports the least. Notably, the year 2020 stands out, revealing a reduction in reported victim counts across all three age groups. This decline is attributed to the impact of COVID-19 and the ensuing lockdown measures, resulting in a seemingly decrease in crime reporting rates.

@@

@fig-totalsum-graph corresponds to the information presented in @tbl-totalsum-table. Visualizing the data by age group across each year reveals that the majority of reported crime victims fall into the "Adult" category. The bars exhibit an ascending trend from 2014 to 2018. However, a conspicuous drop in reported crimes is observed between 2019 and 2020. One plausible explanation for this decline is attributed to the impact of COVID-19, which influenced crime reporting rates during this period.

@@

@fig-r-graph depicts that as of 2018, the youth group of robbery victims diminished and has remained at a lower level. The COVID pandemic is a significant factor influencing the bar graphs depicting the number of robbery victims in different age groups over the years. In 2020, victims of robbery in every age group decreased, indicating a possible impact of lockdown mandates on the decrease in victims.

@@

The onset of the COVID-19 pandemic in late 2019 introduced a noticeable factor that impacted crime trends throughout the subsequent years, especially in 2020 and extending up to 2022. One reasonable explanation is the reduction in physical interactions among individuals, potentially leading to a decline in contact-based offenses such as sexual violence and robbery. However, it is crucial not to overlook the possibility of an increase in domestic violence during this period. Domestic settings may contribute to instances of sexual violence, presenting a contrasting outcome.

@@

The analysis reveals trends in victims of crimes in Toronto from 2014 to 2022. While the overall pattern suggests a decline in reported crimes, the impact of the COVID-19 pandemic is evident in specific crime subtypes and age groups. Sexual violation cases, for instance, witnessed a shift in the dominant age group from adults to youth in 2021. Assault incidents displayed a consistent upward trend, unaffected by the pandemic, but with a reduction in the sizes of child and youth victim groups. In contrast, the number of robbery victims, particularly in the youth group, significantly decreased in 2020, likely influenced by pandemic-related lockdown measures. The findings offer insights between external events and crime statistics.
